---
title: "Spotify Data Analysis"
author: "Khoi Trinh"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analyzing My Spotify Streaming History

All of the codes are based on this [article](https://towardsdatascience.com/explore-your-activity-on-spotify-with-r-and-spotifyr-how-to-analyze-and-visualize-your-stream-dee41cb63526)

First, here are the required library

```{r library, warning=FALSE, message=FALSE}
# REQUIRED LIBRARIES
library(jsonlite)
library(lubridate)
library(gghighlight)
library(spotifyr)
library(tidyverse)
library(knitr)
library(ggplot2)
library(plotly)
```

Let's read in our data, you can find how to get your own spotify data [here](https://www.makeuseof.com/tag/download-privacy-data-spotify/)

```{r read in json}
streamHistory0 <- data.frame(fromJSON("StreamingHistory0.json", flatten = TRUE))
streamHistory1 <- data.frame(fromJSON("StreamingHistory1.json", flatten = TRUE))
streamHistory2 <- data.frame(fromJSON("StreamingHistory2.json", flatten = TRUE))
streamHistory3 <- data.frame(fromJSON("StreamingHistory3.json", flatten = TRUE))
streamHistory4 <- data.frame(fromJSON("StreamingHistory4.json", flatten = TRUE))
streamHistory5 <- data.frame(fromJSON("StreamingHistory5.json", flatten = TRUE))

streamHistory <- Map(c, streamHistory0, streamHistory1, streamHistory2, 
                     streamHistory3, streamHistory4, streamHistory5)
```

On what day did I listen to more or less music?

```{r add date and time}
# Add date and time
mySpotify <- streamHistory %>% 
  as_tibble() %>% 
  mutate_at("endTime", ymd_hm) %>% 
  mutate(endTime = endTime - hours(6)) %>% 
  mutate(date = floor_date(endTime, "day") %>% as_date, 
         seconds = msPlayed / 1000, minutes = seconds / 60)

# Playback activity per week and hours
streamingHours <- mySpotify %>% 
  filter(date >= "2020-01-01") %>% 
  group_by(date) %>% 
  group_by(date = floor_date(date, "week")) %>%
  summarize(hours = sum(minutes) / 60) %>% 
  arrange(date) %>% 
  ggplot(aes(x = date, y = hours)) + 
  geom_col(aes(fill = hours)) +
  scale_fill_gradient(low = "yellow", high = "red") + 
  labs(x= "Date", y= "Hours of music playback") + 
  ggtitle("Weekly Activity On Spoticy")
```

Let's view the plot

```{r streaming plot}
streamingHours
```

What time of the day am I most active?

```{r time of day}
hoursDay <- mySpotify %>% 
  filter(date >= "2019-01-01") %>% 
  group_by(date, hour = hour(endTime), weekday = wday(date, label = TRUE))%>% 
  summarize(minutesListened = sum(minutes))

hoursDay %>% 
  group_by(weekday, hour) %>% 
  summarize(minutes = sum(minutesListened)) %>% 
  ggplot(aes(x = hour, weekday, fill = minutes)) + 
  geom_tile() + 
  scale_fill_gradient(low = "yellow", high = "red") +
  labs(x= "Time of the day", y= "Weekday") + 
  ggtitle("What day and time of day am I most active?")
```

Here is the graph

```{r}
hoursDay
```

How about streaming time by a specific artist?

```{r artist stream, warning=FALSE, message=FALSE}
hoursArtist <- mySpotify %>% 
  group_by(artistName, date = floor_date(date, "month")) %>% 
  summarize(hours = sum(minutes) / 60) %>% 
  ggplot(aes(x = date, y = hours, group = artistName)) + 
  labs(x= "Date", y= "Hours of music playback") + 
  ggtitle("Highlighting Specific Artists") +
  geom_line() + 
  gghighlight(artistName == "Bad Omens" || artistName == "Spiritbox" 
              || artistName == "Linkin Park") 
```

Let's look at the graph

```{r highlighting artist}
hoursArtist
```

Finally, let's get my most listened to artist(s) in 2022

```{r most listened to artsits}
topArtists <- mySpotify %>% 
  filter(date >= "2022-01-01") %>% 
  group_by(artistName) %>% 
  summarize(minutesListened = sum(minutes)) %>% 
  filter(minutesListened >= 1200) %>%
  ggplot(aes(x = artistName, y = minutesListened)) + 
  geom_col(aes(fill = minutesListened)) +
  scale_fill_gradient(low = "yellow", high = "red") + 
  labs(x= "Artist", y= "Minutes of music playback") + 
  ggtitle("My Most Listened To Artists", "> 20 hours listened") +
  theme(axis.text.x = element_text(angle = 90))
```

And the plot

```{r top artists}
topArtists
```